### 2021-01-05 priority-gv.tex
give space to results
Processed minor comments and inserted unresolved comments as todo notes.

We present Priority GV (PGV), a session-typed functional language based on GV~\cite{wadler15,lindleymorris15} which uses priorities \`{a} la Kobayashi~\cite{kobayashi06,padovaninovara15} to enforce deadlock freedom.
Priority GV offers a more fine-grained analysis of communication structures, and by separating channel creation form thread spawning it allows cyclic structures. We start by illustrating PGV with Milner's Cyclic Scheduler~\cite{milner89}.

\begin{example}[Milner's Cyclic Scheduler in PGV]\label{sec:pgv-example}
Suppose a set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, is scheduled to perform some tasks in cyclic order, starting with $\tm{\proc1}$, ending with $\tm{\proc{n}}$, and notifying $\tm{\proc1}$ when all processes have finished.
We implement the scheduler as follows, using $\tm{\prod_{I}\conf{C}_i}$ to denote the parallel composition of the processes $\tm{\conf{C}_i}$, $\tm{i}\in\tm{I}$, and $\tm{\plug{M}{N}}$ to denote the plugging of $\tm{N}$ in the one-hole term-context $\tm{M}$:
\[
  \begin{array}{lrlr}
    \tm{\sched}
    & \defeq & \tm{\res{a_1}{a'_1}{\dots{\res{a_n}{a'_n}{}}}}
               \tm{\res{b_1}{b'_1}{\dots{\res{b_n}{b'_n}{}}}}
               \tm{\res{c_1}{c'_1}{\dots{\res{c_n}{c'_n}{}}}}
               \tm{\res{d}{d'}{}}
    \\ &     & \tm{\begin{array}{lll}
                     (
                     & \phi\;\proc1
                     \parallel
                     \child\;\andthen{\agent1}{\andthen{\wait\;{c'_n}}{\close\;{d}}}
                     \\
                     \parallel
                     &
                     \prod_{2\leq{i}\leq{n}}
                     (\child\;\proc{i} \parallel \child\;\andthen{\wait\;{c'_{i-1}}}{\agent{i}})
                     & )
                   \end{array}}
    \\
    \tm{\agent{i}}
    & \defeq & \tm{\andthen{\close\;{a_i}}{\andthen{\wait\;{b_i}}{\close\;{c_i}}}}
    \\
    \tm{\proc{i}}
    & \defeq & \tm{\andthen{\wait\;{a'_i}}{\plug{M_i}{\andthen{\close\;{b'_i}}{N_i}}}}
  \end{array}
\]
Our scheduler $\tm{\sched}$ consists of set of agents $\tm{\agent{i}}$, $1\leq{i}\leq{n}$, each representing their respective process. Each process $\tm{\proc{i}}$ waits for the signal to start their task on $\tm{a'_i}$, and signals completion on $\tm{b'_i}$. Each agent signals their process to start on $\tm{a_i}$, waits for their process to finish on $\tm{b_i}$, and then signals for the next agent to continue on $\tm{c_i}$. The agent $\tm{\agent0}$ initiates, then waits for every other process to finish, and signals $\tm{\proc1}$ on $\tm{d}$. Every other agent $\tm{\agent{i}}$, $2\leq{i}\leq{n}$ waits on $\tm{c'_{i-1}}$ for the signal to start.
Each of the channels in the scheduler is of a terminated type, and is merely used to synchronise. For simplicity, we let $\tm{\sched}$ be a configuration. However, we could equivalently have represented it as a function which takes three processes as arguments, and creates this configuration.
A~diagram of our scheduler with three processes can be found in \Cref{fig:pgv-scheduler}.
\end{example}

@@

\subsubsection*{Environments}
Typing environments $\ty{\Gamma}$, $\ty{\Delta}$ associate names to types. Environments are linear, so two environments can only be combined as $\ty{\Gamma}, \ty{\Delta}$ if their names are distinct, \ie $\fv(\ty{\Gamma})\cap\fv(\ty{\Delta})=\varnothing$.
\[
\begin{array}{lcl}
  \ty{\Gamma}, \ty{\Delta}
  & \Coloneqq & \ty{\emptyenv}
    \sep        \ty{\Gamma}, \tmty{x}{T}
\end{array}
\]

\subsubsection*{Duality}
Duality plays a crucial role in session types. The two endpoints of a channel are assigned dual types, which ensures that, for instance, whenever one program \emph{sends} a value on a channel, the program on the other end is waiting to \emph{receive}. Each session type $\ty{S}$ has a dual, written $\ty{\co{S}}$. Duality is an involutive function which preserves priorities.
\[
  \ty{\co{\tysend[\cs{o}]{T}{S}}} = \ty{\tyrecv[\cs{o}]{T}{\co{S}}}
  \qquad
  \ty{\co{\tyrecv[\cs{o}]{T}{S}}} = \ty{\tysend[\cs{o}]{T}{\co{S}}}
  \qquad
  \ty{\co{\tyends[\cs{o}]}} = \ty{\co{\tyendr[\cs{o}]}}
  \qquad
  \ty{\co{\tyendr[\cs{o}]}} = \ty{\co{\tyends[\cs{o}]}}
\]

\subsubsection*{Priorities}
Function $\pr(\cdot)$ returns the smallest priority of a session type. The type system guarantees that the top-most connective always holds the smallest priority, so we simply return the priority of the top-most connective:
\[
  \pr(\ty{\tysend[\cs{o}]{T}{S}}) = \cs{o}
  \qquad
  \pr(\ty{\tyrecv[\cs{o}]{T}{S}}) = \cs{o}
  \qquad
  \pr(\ty{\tyends[\cs{o}]})       = \cs{o}
  \qquad
  \pr(\ty{\tyendr[\cs{o}]})       = \cs{o}
\]
We extend the function $\pr(\cdot)$ to types and typing contexts by returning the smallest priority in the type or context, or $\cs{\top}$ if there is no priority. We use $\sqcap$ and $\sqcup$ to denote the minimum and maximum:
\[
\begin{array}{lcl}
  \minpr(\ty{\typrod{T}{U}})                 & = & \minpr({\ty{T}})\sqcap\minpr({\ty{U}}) \\
  \minpr(\ty{\tysum{T}{U}})                  & = & \minpr({\ty{T}})\sqcap\minpr({\ty{U}}) \\
  \minpr(\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}) & = & \cs{p} \\
  \minpr(\ty{\Gamma}, \tmty{x}{A})           & = & \minpr(\ty{\Gamma})\sqcap\pr(\ty{A})
\end{array}
\qquad
\begin{array}{lcl}
  \minpr(\ty{\tyunit})                       & = & \ptop \\
  \minpr(\ty{\tyvoid})                       & = & \ptop \\
  \minpr(\ty{S})                             & = & \pr(\ty{S}) \\
  \minpr(\ty{\emptyenv})                     & = & \ptop
\end{array}
\]

@@

Constants are best understood in conjunction with their typing and reduction rules in~\cref{fig:pgv-typing,fig:pgv-operational-semantics}. Briefly, $\tm{\link}$ links two endpoints together, forwarding messages from one to the other, $\tm{\new}$ creates a new channel and returns a pair of its endpoints, and $\tm{\spawn}$ spawns off its argument as a new thread. The $\tm{\send}$ and $\tm{\recv}$ functions send and receive values on a channel. However, since the typing rules for PGV ensure the linear usage of endpoints, they also return a new copy of the endpoint to continue the session. The $\tm{\close}$ and $\tm{\wait}$ functions close a channel.
We use syntactic sugar to make terms more readable: we write $\tm{\andthen{M}{N}}$ in place of $\tm{\letunit{M}{N}}$, $\tm{\letbind{x}{M}{N}}$ in place of $\tm{(\lambda x.N)\;M}$, and pattern matching functions $\tm{\lambda\unit.M}$ in place of $\tm{\lambda z.\letunit{z}{M}}$ and $\tm{\lambda\pair{x}{y}.M}$ in place of $\tm{\lambda z.\letpair{x}{y}{z}{M}}$.
% We recover $\tm{\fork}$ as $\tm{\lambda x.\letpair{y}{z}{\new\;\unit}{\andthen{\spawn\;{(\lambda\unit.x\;y)}}{z}}}$.

\subsubsection*{Internal and External Choice}
Typically, session-typed languages feature constructs for internal and external choice. In GV, these can be defined in terms of the core language, by sending or receiving a value of a sum type~\cite{lindleymorris15}. We use the following syntactic sugar for internal ($\ty{\tyselect[\cs{o}]{S}{S'}}$) and external ($\ty{\tyoffer[\cs{o}]{S}{S'}}$) choice and their units:
\[
\begin{array}{lcl}
  \ty{\tyselect[\cs{o}]{S}{S'}}
  & \elabarrow & \ty{\tysend[\cs{o}]{(\tysum{\co{S}}{\co{S'}})}{\tyends[\cs{o+1}]}} \\
  \ty{\tyoffer[\cs{o}]{S}{S'}}
  & \elabarrow & \ty{\tyrecv[\cs{o}]{(\tysum{S}{S'})}{\tyendr[\cs{o+1}]}} \\
\end{array}
\qquad
\begin{array}{lcl}
  \ty{\tyselectemp[\cs{o}]}
  & \elabarrow & \ty{\tysend[\cs{o}]{\tyvoid}{\tyends[\cs{o+1}]}} \\
  \ty{\tyofferemp[\cs{o}]}
  & \elabarrow & \ty{\tyrecv[\cs{o}]{\tyvoid}{\tyendr[\cs{o+1}]}} \\
\end{array}
\]
As the syntax for units suggests, these are the binary and nullary forms of the more common n-ary choice constructs $\ty{{\oplus}^{\cs{o}}\{l_i:S_i\}_{i\in{I}}}$ and $\ty{{\with}^{\cs{o}}\{l_i:S_i\}_{i\in{I}}}$, which one may obtain generalising the sum types to variant types. For simplicity, we present only the binary and nullary forms.

@@

While our configuration reduction is based on the standard evaluation for GV, the increased expressiveness of PGV allows us to simplify the relation on two counts.
Firstly, we decompose the $\tm{\fork}$ construct. In GV, $\tm{\fork}$ creates a new channel, spawns a child thread, and, when the child thread finishes, it closes the channel to its parent. In PGV, these are three separate operations: $\tm{\new}$, $\tm{\spawn}$, and $\tm{\close}$. We no longer require that every child thread finishes by returning a terminated channel. Consequently, we also simplify the evaluation of the $\tm{\link}$ construct. In GV, $\tm{\link}$ needs to return a terminated channel after evaluating the link. In PGV, the behaviour matches that of link in CP.
Secondly, our structural congruence is type preserving. Consequently, we can embed it directly into the reduction relation. In GV, this is not the case, and subject reduction relies on proving that if $\equiv\cred$ ends up in an ill-typed configuration, we can rewrite it to a well-typed configuration using $\equiv$.

\subsubsection*{Typing}
\Cref{fig:pgv-typing} gives the typing rules for PGV.
Typing rules for terms are at the top of \cref{fig:pgv-typing}. Terms are typed by a judgement $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ stating that ``a term $\tm{M}$ has type $\ty{T}$ and an upper bound on its priority $\cs{p}$ under the typing environment $\ty{\Gamma}$''. Typing for the linear $\lambda$-calculus is standard. Linearity is ensured by splitting environments on branching rules, requiring that the environment in the variable rule consists of just the variable, and the environment in the constant and unit rules are empty. Constants $\tm{K}$ are typed using type schemas, and embedded using \LabTirName{T-Const} (mid of \cref{fig:pgv-typing}).

@@

We must show that substitution preserves priority constraints. For this, we prove~\cref{lem:pgv-value-done}, which shows that values have finished all their communication, and that any priorities in the type of the value come from the typing environment.
\begin{compacttheorems}
  \begin{restatablelemma}{lempgvvaluedone}
    \label{lem:pgv-value-done}
    If $\tseq[\cs{p}]{\ty{\Gamma}}{V}{T}$, then $\cs{p}=\cs{\pbot}$, and $\minpr(\ty{\Gamma})=\minpr(\ty{T})$.
  \end{restatablelemma}
  \begin{proof}
    By induction on the derivation of $\tseq[\cs{p}]{\ty{\Gamma}}{V}{T}$. We demonstrate the case for \LabTirName{T-Inl}, as it illustrates the need for the side conditions on \LabTirName{T-Inl} and \LabTirName{T-Inr}:
    \begin{case*}[\LabTirName{T-Inl}]
      The induction hypothesis gives us $\cs{p}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr{(\ty{T})}$. We know $\pr(\ty{T})=\pr({\ty{U}})$, hence $\pr(\ty{\Gamma})=\pr(\ty{\tysum{T}{U}})$.
      \begin{mathpar}
        \small
        \inferrule*{
          \tseq[\cs{p}]{\ty{\Gamma}}{V}{T}
          \\
          \pr(\ty{T})=\pr(\ty{U})
        }{\tseq[\cs{p}]{\ty{\Gamma}}{\inl{V}}{\tysum{T}{U}}}
      \end{mathpar}
    \end{case*}
    The full proof can be found in~\cref{prf:lem-pgv-value-done}.
  \end{proof}
  \begin{restatablelemma}{lempgvsubstitution}[Substitution]
    \label{lem:pgv-substitution}
    \hfill\\%newline before theorem statement
    If $\tseq[\cs{p}]{\ty{\Gamma},\tmty{x}{U'}}{M}{T}$ and $\tseq[\cs{q}]{\ty{\Theta}}{V}{U'}$, then $\tseq[\cs{p}]{\ty{\Gamma},\ty{\Theta}}{\subst{M}{V}{x}}{T}$.
  \end{restatablelemma}
  \begin{proof}
    By induction on the derivation of $\tseq[\cs{p}]{\ty{\Gamma},\tmty{x}{U'}}{M}{T}$. Whenever the contexts in the premises and conclusion differ, we use \cref{lem:pgv-value-done} to prove that the priorities are preserved. We demonstrate the case for \LabTirName{T-Lam}:
    \begin{case*}[\LabTirName{T-Lam}]
      By \cref{lem:pgv-value-done}, $\pr(\ty{\Theta})=\pr(\ty{U'})$, hence $\pr(\ty{\Gamma},\ty{\Theta})=\pr(\ty{\Gamma},\ty{U'})$.
      \begin{mathpar}
        \small
        \inferrule*{
          \tseq[\cs{q}]{\ty{\Gamma},\tmty{x}{U'},\tmty{y}{T}}{M}{U}
        }{\tseq[\cs{\pbot}]{\ty{\Gamma},\tmty{x}{U'}}
          {\lambda y.M}
          {\tylolli[\cs{\pr(\ty{\Gamma},\ty{U'})},\cs{q}]{T}{U}}}
        \substarrow{V}{x}
        \inferrule*{
          \tseq[\cs{q}]{\ty{\Gamma},\ty{\Theta},\tmty{y}{T}}{\subst{M}{V}{x}}{U}
        }{\tseq[\cs{\pbot}]{\ty{\Gamma},\ty{\Theta}}
          {\lambda y.\subst{M}{V}{x}}
          {\tylolli[\cs{\pr(\ty{\Gamma},\ty{\Theta})},\cs{q}]{T}{U}}}
      \end{mathpar}
    \end{case*}
    The complete proof can be found in~\cref{prf:lem-pgv-substitution}.
  \end{proof}
\end{compacttheorems}
The proofs for subject congruence and subject reduction are standard inductions. The details can be found in \cref{prf:lem-pgv-subject-reduction-terms,prf:lem-pgv-subject-congruence,prf:thm-pgv-subject-reduction-confs}.
\begin{compacttheorems}
  \begin{restatablelemma}{lempgvsubjectreductionterms}[Subject Reduction, $\tred$]
    \label{lem:pgv-subject-reduction-terms}
    \hfill\\%newline before theorem statement
    If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ and $\tm{M}\tred\tm{M'}$,
    then $\tseq[\cs{p}]{\ty{\Gamma}}{M'}{T}$.
  \end{restatablelemma}
  \begin{restatablelemma}{lempgvsubjectcongruence}[Subject Congruence, $\equiv$]
    \label{lem:pgv-subject-congruence}
    \hfill\\%newline before theorem statement
    If $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$ and $\tm{\conf{C}}\equiv\tm{\conf{C'}}$,
    then $\cseq[\phi]{\ty{\Gamma}}{\conf{C'}}$.
  \end{restatablelemma}
  \begin{restatabletheorem}{thmpgvsubjectreductionconfs}[Subject Reduction, $\cred$]
    \label{thm:pgv-subject-reduction-confs}
    \hfill\\%newline before theorem statement
    If $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$ and $\tm{\conf{C}}\cred\tm{\conf{C'}}$,
    then $\cseq[\phi]{\ty{\Gamma}}{\conf{C'}}$.
  \end{restatabletheorem}
\end{compacttheorems}

>>>>

We present Priority GV (PGV), a session-typed functional language based on GV~\cite{wadler15,lindleymorris15} which uses priorities \`{a} la Padovani~\cite{kobayashi06,padovaninovara15} to enforce deadlock freedom.
Priority GV offers a more fine-grained analysis of communication structures, and by separating channel creation form thread spawning it allows cyclic structures. We start by illustrating PGV with Milner's Cyclic Scheduler~\cite{milner89}.

\todo{The $\phi$ and $\child$ symbols are not explained at this stage and become clear only later. Either warn the reader or remove them altogether.}

\todo{An explanation is needed about what M and N represents and how the hole-construction models this "signal b when finished" behaviour.}

\todo{It is difficult to understand how a "close d" operation is supposed to work inside a (nu d d') construct without a corresponding wait d.}

\begin{example}[Milner's Cyclic Scheduler in PGV]\label{sec:pgv-example}
Suppose a set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, is scheduled to perform some tasks in cyclic order, starting with $\tm{\proc1}$, ending with $\tm{\proc{n}}$, and notifying $\tm{\proc1}$ when all processes have finished.
We implement the scheduler as follows, using $\tm{\prod_{I}\conf{C}_i}$ to denote the parallel composition of the processes $\tm{\conf{C}_i}$, $\tm{i}\in\tm{I}$, and $\tm{\plug{M}{N}}$ to denote the plugging of $\tm{N}$ in the one-hole term-context $\tm{M}$:
\[
  \begin{array}{lrlr}
    \tm{\sched}
    & \defeq & \tm{\res{a_1}{a'_1}{\dots{\res{a_n}{a'_n}{}}}}
               \tm{\res{b_1}{b'_1}{\dots{\res{b_n}{b'_n}{}}}}
               \tm{\res{c_1}{c'_1}{\dots{\res{c_n}{c'_n}{}}}}
               \tm{\res{d}{d'}{}}
    \\ &     & \tm{\begin{array}{lll}
                     (
                     & \phi\;\proc1
                     \parallel
                     \child\;\andthen{\agent1}{\andthen{\wait\;{c'_n}}{\close\;{d}}}
                     \\
                     \parallel
                     &
                     \prod_{2\leq{i}\leq{n}}
                     (\child\;\proc{i} \parallel \child\;\andthen{\wait\;{c'_{i-1}}}{\agent{i}})
                     & )
                   \end{array}}
    \\
    \tm{\agent{i}}
    & \defeq & \tm{\andthen{\close\;{a_i}}{\andthen{\wait\;{b_i}}{\close\;{c_i}}}}
    \\
    \tm{\proc{i}}
    & \defeq & \tm{\andthen{\wait\;{a'_i}}{\plug{M_i}{\andthen{\close\;{b'_i}}{N_i}}}}
  \end{array}
\]
Our scheduler $\tm{\sched}$ consists of set of agents $\tm{\agent{i}}$, $1\leq{i}\leq{n}$, each representing their respective process. Each process $\tm{\proc{i}}$ waits for the signal to start their task on $\tm{a'_i}$, and signals completion on $\tm{b'_i}$. Each agent signals their process to start on $\tm{a_i}$, waits for their process to finish on $\tm{b_i}$, and then signals for the next agent to continue on $\tm{c_i}$. The agent $\tm{\agent1}$ initiates, then waits for every other process to finish, and signals $\tm{\proc1}$ on $\tm{d}$. Every other agent $\tm{\agent{i}}$, $2\leq{i}\leq{n}$ waits on $\tm{c'_{i-1}}$ for the signal to start.
Each of the channels in the scheduler is of a terminated type, and is merely used to synchronise. For simplicity, we let $\tm{\sched}$ be a configuration.
A~diagram of our scheduler with three processes can be found in \Cref{fig:pgv-scheduler}.
\end{example}

@@

\subsubsection*{Environments}
Typing environments $\ty{\Gamma}$, $\ty{\Delta}$ associate types to names. Environments are linear, so two environments can only be combined as $\ty{\Gamma}, \ty{\Delta}$ if their names are distinct, \ie $\fv(\ty{\Gamma})\cap\fv(\ty{\Delta})=\varnothing$.
\[
\begin{array}{lcl}
  \ty{\Gamma}, \ty{\Delta}
  & \Coloneqq & \ty{\emptyenv}
    \sep        \ty{\Gamma}, \tmty{x}{T}
\end{array}
\]

\subsubsection*{Duality}
Duality plays a crucial role in session types. The two endpoints of a channel are assigned dual types, ensuring that, for instance, whenever one program \emph{sends} a value on a channel, the program on the other end is waiting to \emph{receive}. Each session type $\ty{S}$ has a dual, written $\ty{\co{S}}$. Duality is an involutive function which preserves priorities.
\[
  \ty{\co{\tysend[\cs{o}]{T}{S}}} = \ty{\tyrecv[\cs{o}]{T}{\co{S}}}
  \qquad
  \ty{\co{\tyrecv[\cs{o}]{T}{S}}} = \ty{\tysend[\cs{o}]{T}{\co{S}}}
  \qquad
  \ty{\co{\tyends[\cs{o}]}} = \ty{\co{\tyendr[\cs{o}]}}
  \qquad
  \ty{\co{\tyendr[\cs{o}]}} = \ty{\co{\tyends[\cs{o}]}}
\]

\subsubsection*{Priorities}
Function $\pr(\cdot)$ returns the smallest priority of a session type. The type system guarantees that the top-most connective always holds the smallest priority, so we simply return the priority of the top-most connective:
\[
  \pr(\ty{\tysend[\cs{o}]{T}{S}}) = \cs{o}
  \qquad
  \pr(\ty{\tyrecv[\cs{o}]{T}{S}}) = \cs{o}
  \qquad
  \pr(\ty{\tyends[\cs{o}]})       = \cs{o}
  \qquad
  \pr(\ty{\tyendr[\cs{o}]})       = \cs{o}
\]
We extend the function $\pr(\cdot)$ to types and typing contexts by returning the smallest priority in the type or context, or $\cs{\top}$ if there is no priority. We use $\sqcap$ and $\sqcup$ to denote the minimum and maximum:
\[
\begin{array}{lcl}
  \minpr(\ty{\typrod{T}{U}})                 & = & \minpr({\ty{T}})\sqcap\minpr({\ty{U}}) \\
  \minpr(\ty{\tysum{T}{U}})                  & = & \minpr({\ty{T}})\sqcap\minpr({\ty{U}}) \\
  \minpr(\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}) & = & \cs{p} \\
  \minpr(\ty{\Gamma}, \tmty{x}{A})           & = & \minpr(\ty{\Gamma})\sqcap\minpr(\ty{A})
\end{array}
\qquad
\begin{array}{lcl}
  \minpr(\ty{\tyunit})                       & = & \ptop \\
  \minpr(\ty{\tyvoid})                       & = & \ptop \\
  \minpr(\ty{S})                             & = & \pr(\ty{S}) \\
  \minpr(\ty{\emptyenv})                     & = & \ptop
\end{array}
\]

@@

Constants are best understood in conjunction with their typing and reduction rules in~\cref{fig:pgv-typing,fig:pgv-operational-semantics}. Briefly, $\tm{\link}$ links two endpoints together, forwarding messages from one to the other, $\tm{\new}$ creates a new channel and returns a pair of its endpoints, and $\tm{\spawn}$ spawns off its argument as a new thread. The $\tm{\send}$ and $\tm{\recv}$ functions send and receive values on a channel. However, since the typing rules for PGV ensure the linear usage of endpoints, they also return a new copy of the endpoint to continue the session. The $\tm{\close}$ and $\tm{\wait}$ functions close a channel.
We use syntactic sugar to make terms more readable: we write $\tm{\andthen{M}{N}}$ in place of $\tm{\letunit{M}{N}}$, $\tm{\letbind{x}{M}{N}}$ in place of $\tm{(\lambda x.N)\;M}$, and pattern matching functions $\tm{\lambda\unit.M}$ in place of $\tm{\lambda z.\letunit{z}{M}}$ and $\tm{\lambda\pair{x}{y}.M}$ in place of $\tm{\lambda z.\letpair{x}{y}{z}{M}}$.
\todo{Confusion arises from the difference between what is considered a native operator and a shortcut. For instance, "unit let" and "pair let" are native (representing sequence and new channel decomposition) while "singleton let" is a shortcut. The initial GV syntax was clear, this one is a bit confusing and might need some cleaning. At least explain why M;N cannot be a primitive construct.}
% We recover $\tm{\fork}$ as $\tm{\lambda x.\letpair{y}{z}{\new\;\unit}{\andthen{\spawn\;{(\lambda\unit.x\;y)}}{z}}}$.

\subsubsection*{Internal and External Choice}
\todo{You might want to clarify (repeat) that external choice can be modeled by sum type communication only because the language uses linear channels.}
\todo{We also might want to emphasise that the modelling of choice in GV by~\cite{lindleymorris15} is unsound.}
Typically, session-typed languages feature constructs for internal and external choice. In GV, these can be defined in terms of the core language, by sending or receiving a value of a sum type~\cite{lindleymorris15}. We use the following syntactic sugar for internal ($\ty{\tyselect[\cs{o}]{S}{S'}}$) and external ($\ty{\tyoffer[\cs{o}]{S}{S'}}$) choice and their units:
\[
\begin{array}{lcl}
  \ty{\tyselect[\cs{o}]{S}{S'}}
  & \elabarrow & \ty{\tysend[\cs{o}]{(\tysum{\co{S}}{\co{S'}})}{\tyends[\cs{o+1}]}} \\
  \ty{\tyoffer[\cs{o}]{S}{S'}}
  & \elabarrow & \ty{\tyrecv[\cs{o}]{(\tysum{S}{S'})}{\tyendr[\cs{o+1}]}} \\
\end{array}
\qquad
\begin{array}{lcl}
  \ty{\tyselectemp[\cs{o}]}
  & \elabarrow & \ty{\tysend[\cs{o}]{\tyvoid}{\tyends[\cs{o+1}]}} \\
  \ty{\tyofferemp[\cs{o}]}
  & \elabarrow & \ty{\tyrecv[\cs{o}]{\tyvoid}{\tyendr[\cs{o+1}]}} \\
\end{array}
\]
As the syntax for units suggests, these are the binary and nullary forms of the more common n-ary choice constructs $\ty{{\oplus}^{\cs{o}}\{l_i:S_i\}_{i\in{I}}}$ and $\ty{{\with}^{\cs{o}}\{l_i:S_i\}_{i\in{I}}}$, which one may obtain generalising the sum types to variant types. For simplicity, we present only the binary and nullary forms.

@@

While our configuration reduction is based on the standard evaluation for GV, the increased expressiveness of PGV allows us to simplify the relation on two counts.
Firstly, we decompose the $\tm{\fork}$ construct. In GV, $\tm{\fork}$ creates a new channel, spawns a child thread, and, when the child thread finishes, it closes the channel to its parent. In PGV, these are three separate operations: $\tm{\new}$, $\tm{\spawn}$, and $\tm{\close}$. We no longer require that every child thread finishes by returning a terminated channel. Consequently, we also simplify the evaluation of the $\tm{\link}$ construct. In GV, $\tm{\link}$ needs to return a terminated channel after evaluating the link. In PGV, the behaviour matches that of link in CP.\todo{Please clarify in what way it matches the behaviour of link in CP.}
Secondly, our structural congruence is type preserving. Consequently, we can embed it directly into the reduction relation. In GV, this is not the case, and subject reduction relies on proving that if $\equiv\cred$ ends up in an ill-typed configuration, we can rewrite it to a well-typed configuration using $\equiv$.

\subsubsection*{Typing}
\todo{This is a type-and-effect system and should probably be presented as such, as in [37].}
\Cref{fig:pgv-typing} gives the typing rules for PGV.
Typing rules for terms are at the top of \cref{fig:pgv-typing}. Terms are typed by a judgement $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ stating that ``a term $\tm{M}$ has type $\ty{T}$ and an upper bound on its priority $\cs{p}$ under the typing environment $\ty{\Gamma}$''. Typing for the linear $\lambda$-calculus is standard. Linearity is ensured by splitting environments on branching rules, requiring that the environment in the variable rule consists of just the variable, and the environment in the constant and unit rules are empty. Constants $\tm{K}$ are typed using type schemas, and embedded using \LabTirName{T-Const} (mid of \cref{fig:pgv-typing}).

\todo{From the way the typing rules \LabTirName{T-Var} and \LabTirName{T-Const} are formulated, it is clear that all variables are treated as linear resources, even if they happen to have type 1 or another non-linear type. This could be worth mentioning in the text, saying that the type system can be easily extended to ignore non-linear resources.}

@@

We must show that substitution preserves priority constraints. For this, we prove~\cref{lem:pgv-value-done}, which shows that values have finished all their communication, and that any priorities in the type of the value come from the typing environment.
\begin{compacttheorems}
  \begin{restatablelemma}{lempgvvaluedone}
    \label{lem:pgv-value-done}
    If $\tseq[\cs{p}]{\ty{\Gamma}}{V}{T}$, then $\cs{p}=\cs{\pbot}$, and $\minpr(\ty{\Gamma})=\minpr(\ty{T})$.
  \end{restatablelemma}
  \begin{restatablelemma}{lempgvsubstitution}[Substitution]
    \label{lem:pgv-substitution}
    \hfill\\%newline before theorem statement
    If $\tseq[\cs{p}]{\ty{\Gamma},\tmty{x}{U'}}{M}{T}$ and $\tseq[\cs{q}]{\ty{\Theta}}{V}{U'}$, then $\tseq[\cs{p}]{\ty{\Gamma},\ty{\Theta}}{\subst{M}{V}{x}}{T}$.
  \end{restatablelemma}
  \begin{restatablelemma}{lempgvsubjectreductionterms}[Subject Reduction, $\tred$]
    \label{lem:pgv-subject-reduction-terms}
    \hfill\\%newline before theorem statement
    If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ and $\tm{M}\tred\tm{M'}$,
    then $\tseq[\cs{p}]{\ty{\Gamma}}{M'}{T}$.
  \end{restatablelemma}
  \begin{restatablelemma}{lempgvsubjectcongruence}[Subject Congruence, $\equiv$]
    \label{lem:pgv-subject-congruence}
    \hfill\\%newline before theorem statement
    If $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$ and $\tm{\conf{C}}\equiv\tm{\conf{C'}}$,
    then $\cseq[\phi]{\ty{\Gamma}}{\conf{C'}}$.
  \end{restatablelemma}
  \begin{restatabletheorem}{thmpgvsubjectreductionconfs}[Subject Reduction, $\cred$]
    \label{thm:pgv-subject-reduction-confs}
    \hfill\\%newline before theorem statement
    If $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$ and $\tm{\conf{C}}\cred\tm{\conf{C'}}$,
    then $\cseq[\phi]{\ty{\Gamma}}{\conf{C'}}$.
  \end{restatabletheorem}
\end{compacttheorems}
The proofs are standard inductions. The details can be found in \cref{prf:lem-pgv-value-done,prf:lem-pgv-substitution,prf:lem-pgv-subject-reduction-terms,prf:lem-pgv-subject-congruence,prf:thm-pgv-subject-reduction-confs}.
